\textcolor{blue}{Supplementary information for In-silico molecular transport via perivascular networks in the human intracranial space}

### 2025-06-26 paper/supplementary.tex
Updates in the supplementary

\renewcommand\thefigure{S\arabic{figure}}    
\renewcommand\thesection{S\arabic{section}}

@@

These sections provide more detailed descriptions of the mathematical models and numerical approximations considered.
We solve the models presented in Methods numerically
using finite element methods. The following subsections provide more
details, model-by-model.

\subsection{Nomenclature and notation} We begin by including \Cref{table:geometry_notation}, \Cref{table:variable}, and \Cref{table:mesh}, which summarize the common notation used for the geometry, model variables and parameters, and finite element discretization, respectively.
\begin{table}[h!] 
\centering
\caption{Geometrical notation, refer to (Fig.~1A, Fig.~2A--C)}
\label{table:geometry_notation}
\begin{tabular}{ll}
\hline
\textbf{Symbol} & \textbf{Description} \\
\hline
$\Omega$ & Full intracranial domain: $\Omega = \Omega_{\mathrm{PAR}} \cup \Omega_{\mathrm{CSF}}$ \\
$\Omega_{\mathrm{PAR}}$ & Brain parenchyma domain \\
$\Omega_{\mathrm{CSF}}$ & Cerebrospinal‐fluid (CSF) domain \\
$\partial\Omega_{\mathrm{PAR}}\cap\partial\Omega_{\mathrm{CSF}}$ & Parenchyma–CSF interface: $\partial \Omega_{\rm CSF} \bigcap \partial
\Omega_{\rm PAR} = \Gamma_{\rm LV} \cup \Gamma_{\rm pia}$ \\
$\Gamma_{\mathrm{LV}}$ & Lateral‐ventricle surface (part of parenchyma–CSF interface) \\
$\Gamma_{\mathrm{pia}}$ & Pial surface (remainder of parenchyma–CSF interface) \\
$\Gamma_{\mathrm{SSAS}}$ & Lower CSF boundary toward spinal subarachnoid space \\
$\Gamma_{\mathrm{AM}}$ & Outer CSF boundary toward arachnoid/dura \\
$\Gamma_{\mathrm{AM\text{-}L}}$, $\Gamma_{\mathrm{AM\text{-}U}}$ & Lower/upper parts of $\Gamma_{\mathrm{AM}}$ \\
$\Gamma_{\mathrm{SC}}$ & Parenchyma–spinal‐cord boundary: $\partial\Omega_{\mathrm{PAR}}\setminus\partial\Omega_{\mathrm{CSF}}$ \\
$\Lambda_{a},\Lambda_{v}$ & Arterial/venous PVS centerline networks \\
$\Lambda_{i}$ & Individual PVS centerline segment ($\Lambda_a^i$, $\Lambda_v^i$ corresponding to arterial and venous segments) \\
$L_i$ & Length of a PVS segment $\Lambda_i$ \\
$\Omega_{i}$ & Perivascular space around $\Lambda_{i}$ \\
$R_{1}^{i},\,R_{2}^{i}$ & Inner/outer radii of the annular PVS $\Omega_{i}$ \\
$\Gamma_{a},\,\Gamma_{v}$ & Outer lateral surfaces of periarterial/perivenous PVSs \\
$P$ & Perimeter of the PVS cross‐section \\
$\mathcal V$ & Graph nodes of the PVS network \\
$\mathcal B\subset  \mathcal V$ & Internal (bifurcation) nodes \\
$\mathcal N\subset  \mathcal V$ & End (leaf) nodes \\
\hline
\end{tabular}
\end{table}

\begin{table}[htb!]
\centering 
\caption{Variable and parameter notation}
\label{table:variable}
\begin{tabular}{ll}
\hline
\textbf{Symbol} & \textbf{Description} \\
\hline
$\phi$ & Fluid volume fraction (porosity) \\
$D$ & Effective diffusion coefficient ($D_{\mathrm{PAR}} = D^*$ or $D_{\mathrm{CSF}} = D(1+R)$ \\
$\mathbf{u}$ & Convective velocity ($\mathbf{u}_{\mathrm{PAR}} = 0$ or $\mathbf{u}_{\mathrm{CSF}}$ computed from steady state flow) \\
$c$ & Solute concentration in 3D domains \\
$\bar c$ & Lateral average of $c$ on the PVS boundary \\
$\xi$ & Permeability coefficient between the periarterial and perivenous networks
and their surroundings\\
$\delta_{\Gamma}$ & Dirac term on the PVS outer surfaces \\
$\beta_{\mathrm{pia}}$ & Permeability of pial/ventricle interface \\
$\beta_{\mathrm{exit}}$ & Efflux rate on $\Gamma_{\mathrm{AM\text{-}U}}$ \\
$g_{\mathrm{influx}}$ & Influx on $\Gamma_{\mathrm{SSAS}}$ \\
$A(s)$ & Cross‐sectional area of PVS at $s$ \\
$\hat D$ & 1D effective diffusion in PVS \\
$\hat u$ & 1D convective velocity in PVS \\
$p$ & Pressure in CSF (Stokes) or PVS (network) \\
$\hat{q}$ & Volumetric flux in PVS network \\
$\kappa$ & Hydraulic conductivity in annular PVS \\
$\mu$ & Dynamic viscosity of CSF \\
$R_{\rm CSF, 0}$ & Resistance parameter for CSF efflux \\
$u_{\mathrm{in}}$ & Imposed fluid influx on $\Gamma_{\mathrm{LV}}$ \\
$\omega,\;\rho$ & Angular frequency, CSF density (for dispersion estimates) \\
$\alpha_c,\, \alpha_r$ & Womersley number (cardiac, respiratory) \\
$h_{\rm SAS}$ & Mean width of the subarachnoid space (for dispersion estimates) \\
$R_c,\,R_r$ & Spatially-varying dispersion enhancement factors (cardiac, respiratory) \\

\hline
\end{tabular}
\end{table}

\begin{table}[h!]
\centering
\caption{Common mesh and finite‐element notation}
\label{table:mesh}
\begin{tabular}{ll}
\hline
\textbf{Symbol} & \textbf{Description} \\
\hline
$\mathcal{T}$ & Tetrahedral mesh of $\Omega$ \\
$\mathcal{T}_{\mathrm{PAR}},\,\mathcal{T}_{\mathrm{CSF}}$ & Restrictions of $\mathcal T$ to $\Omega_{\mathrm{PAR}},\Omega_{\mathrm{CSF}}$ \\
$E$ & Tetrahedral cell in $\mathcal T$ \\
$\mathcal F_{i}$ & Interior facets of $\mathcal T$ \\
$\mathcal F_{i,\mathrm{PAR}},\, \mathcal F_{i,\mathrm{CSF}}$ & Interior facets in $\Omega_{\mathrm{PAR}}$, $\Omega_{\mathrm{CSF}}$ \\
$\mathcal F_{e}$ & Exterior facets of CSF mesh ($\mathcal F_{\mathrm{pia}}\cup \mathcal F_{\mathrm{LV}}\cup \mathcal F_{\mathrm{AM\text{-}L}}\cup \mathcal \mathcal F_{\mathrm{AM\text{-}U}}\cup \mathcal F_{\mathrm{SSAS}}$) \\
$\mathcal F _{\mathrm{pia}},\,\mathcal F_{\mathrm{LV}},\, \mathcal F_{\mathrm{AM\text{-}L}},\,\mathcal F_{\mathrm{AM\text{-}U}},\, \mathcal F_{\mathrm{SSAS}}$ & Facets on corresponding boundaries \\
$V_{h}$ & Discontinuous Galerkin space on $\Omega$ of degree $1$\\
$\widehat V_{h}$ & Continuous Galerkin space on $\Lambda$ of degree $1$.  \\
% $\widehat V_{h,0}$, $\widehat V_{h,g}$ & Continuous Galerkin space on network with homogeneous or prescribed BCs \\
$\mathrm{BDM}_{2}$ & Brezzi–Douglas–Marini space of degree 2 \\
$Q_{h}$ & Pressure space on $\mathcal T_{\mathrm{CSF}}$  \\
$\mathcal{I}_a, \mathcal{I}_v$ & 1D meshes of arterial/venous networks $\Lambda_{a},\Lambda_{v}$ \\
$\widehat{Z}_{h}(\mathcal{I})$ & Space of piecewise constants ($P^0$) on network mesh $I$ \\
$\tau$ & Time step size \\
\hline
\end{tabular}
\end{table}

@@

In addition, we consider two sets of perivascular networks: a
periarterial network $\Lambda_a$ represented by the (connected)
centerlines $\Lambda^i_{a}$ of the arterial tree, and a perivenous
network $\Lambda_v$ associated with the centerlines $\Lambda^i_{v}$
of the veins. Recall that we represent the vascular domains as the
union of cylindrical vessels of radius $R_1^i$ surrounding the
centerlines $\Lambda^i$. Moreover, we consider the surrounding
perivascular spaces as the union of annular cylinders $\Omega^i$ of
inner radius $R_1^i$ and outer radius $R_2^i > R_1^i$ and thus of
width $R_2^i - R_1^i$, and remark that we interpret $R^i_2$ and $R^i_1$ as temporal averages (fixed in time), as the pulsations considered in \Cref{sec:app:peristalsis} are beyond the temporal resolution of our model. We denote the outer lateral surface of the
periarterial and perivenous spaces by $\Gamma_a$ and $\Gamma_v$,
respectively. We omit the subscript $a$ or $v$ when referring to any
such network, vessel segment or perivascular outer surface. We assume
that $\Lambda$ is parametrized by the coordinate $s$, and with a minor
abuse of notation, simply write $s \in \Lambda$ to represent the point
${\bm \lambda}(s)$ on $\Lambda$ corresponding to $s$. Finally, we view
each perivascular network both as a geometric domain and as a directed
graph with the centerlines $\{ \Lambda^i \}$ as oriented edges and the
connections as nodes $\mathcal{V}$.

@@

\textit{Finite element solution of the incompressible Stokes equations}.
We consider an H(div)-based finite element approximation of the incompressible Stokes equations~\eqref{eq:stokes} defined over $\Omega_{\rm CSF}$ with the boundary conditions~\eqref{eq:stokes_bcs}. Following\cite{hong2016robust}, we approximate the velocity field $\bm u$ and the pressure field $p$ with the following finite element spaces:
\begin{align*}
  \bm V_{h,g} &= \{ \bm v  \in H(\mathrm{div}, \Omega_{\mathrm{CSF}}): \;
  \bm v \vert_{E} \in \mathrm{BDM}^2(E), \; E \in \mathcal{T}_{\rm CSF}; \;
  \bm v \cdot \bm n = 0
  \text{ on } \Gamma_{\rm{pia}} \cup \Gamma_{\rm AM-L} \cup \Gamma_{\rm{SSAS}}, \;\; \bm v \cdot \bm n =  g %\frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}} 
  \text{ on } \Gamma_{\mathrm{LV}} \}  \\ 
  Q_h  &= \{q \in L^2(\Omega_{\rm CSF}): \; q \vert_E \in \mathrm{P}^1(E),
  \; E \in \mathcal{T}_{\rm CSF} \}. 
\end{align*}
Here, $H(\mathrm{div}, \Omega_{\rm{CSF}})$ is the space of $L^2(\Omega_{\rm{CSF}})$ vector fields with $L^2(\Omega_{\rm{CSF}})$ divergence, $\mathrm{BDM}^2$ is the Brezzi-Douglas--Marini element~\cite{brezzi1987mixed} of degree 2, $g$ is a given constant, and $\bm n$ is the unit outward normal vector to each facet. Given any vector $\bm v$, the normal and tangential components on each facet are denoted and given by 
$$ \bm v_n = (\bm v \cdot \bm n) \bm n, \quad \bm v_t = \bm v - \bm v_n. 
$$ 
Since $\bm V_{h,g} \subset H(\mathrm{div}, \Omega_{\rm CSF})$, then $[\bm
  v_n] = 0 $ on $\mathcal{F}_{i, \rm CSF}$, the interior facets to $\Omega_{\mathrm{CSF}}$. Continuity in the
tangential component is enforced weakly via interior penalization. For
convenience, we collect all facets exterior to the CSF space  $\Gamma_{\rm AM-U}$ in the set
$$\mathcal{F}_{e} = \mathcal{F}_{\rm pia} \cup \mathcal{F}_{\rm LV}
\cup \mathcal{F}_{\rm AM-L}\cup \mathcal{F}_{\rm AM-U}  \cup \mathcal{F}_{\rm SSAS}, $$ 
where facets
lying on the pial interface $\Gamma_{\rm pia}$ are denoted by
$\mathcal{F}_{\rm pia}$, on the lower and upper outer (arachnoid) boundary
$\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$ by $\mathcal{F}_{\rm AM-L}$ and $\mathcal{F}_{\rm AM-U}$, respectively, 
%on the boundary toward the spinal cord $\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$
 on the boundary towards the spinal SAS $\Gamma_{\rm SSAS}$ by
$\mathcal{F}_{\rm SSAS}$, and on the surface of the lateral ventricles by $\mathcal{F}_{\rm LV}$. Now, define the form 
\begin{multline}
  \mathcal{A}_h(\bm u, \bm v)
  = \sum_{E \in \mathcal{T}_{\rm CSF}} \int_{E} \mu \nabla \bm u : \nabla \bm v
 + \sum_{F \in \mathcal{F}_{i, \rm CSF} \cup \mathcal{F}_{e}} \left (
  -\int_{F} \mu \{\nabla \bm u\}  \bm n_F \cdot [\bm v_t] 
  - %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
  \int_{F} \mu \{\nabla \bm v\}  \bm n_F \cdot [\bm u_t]
  + %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
\int_{F}   \frac{\sigma \mu}{h_F} [\bm u_t ] \cdot [\bm v_t] \right ),   
\end{multline}
where on exterior facets the average and jump operators take the one-sided values. We set the penalty parameter for the tangential continuity to be $\sigma = 20$. The finite element discretization of the incompressible Stokes equations is then to find $(\bm u_h, p_h) \in \bm V_{h,g} \times Q_h$ with $g  = \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}}$  such that 
\begin{subequations}
\begin{align}
  \mathcal{A}_h(\bm u_h, \bm v_h)
  + \sum_{F \in \mathcal{F}_{\rm AM-U}} \int_{F} R_{\rm CSF, 0} (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n)
  - \int_{\Omega_{\rm CSF}} \nabla \cdot \bm v_h \, p_h  &= 0 \qquad \foralls \bm v_h \in \bm V_{h,0} \\ 
\int_{\Omega_{\rm CSF}} \nabla \cdot \bm u_h \, q_h  &= 0 \qquad \foralls q_h \in Q_h.
\end{align}
\end{subequations}

@@

\textit{Finite element solution of the perivascular network equations}. 
We consider meshes $\mathcal{I}_a, \mathcal{I}_v$ representing a
conforming subdivision of each of the perivascular networks
$\Lambda_a$, $\Lambda_v$. Relative to each mesh $\mathcal{I}$, we define the
space of (discontinuous) piecewise constants $\widehat{Z}_h(\mathcal{I})$ and define
spaces of continuous piecewise linears $\hat{V}_{h, g}(\mathcal{I})$ with
prescribed boundary node values (on $\mathcal{N}$) given by $g$:
\begin{align}
  \hat{Z}_h(\mathcal{I}) &= \{z \in L^2(\Lambda) ; \, z \vert_{\Lambda_i} \in \mathrm{P}^{0}(\Lambda_i) \, \foralls \Lambda^i \in \mathcal{I}\}, \\
  \hat{V}_{h, g}(\mathcal{I}) &= \{ v \in C^0(\Lambda) ; \, v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda_i), \, v(x) = g(x) \, \foralls x \in \mathcal{N} \} .
  %Q_h^{\rm PVS} &= \{ q_h \in C^0(\Lambda),  q_h \vert_{\Lambda^i} \in \mathrm{P}^k(\Lambda_i), q_h \vert_{\Gamma_{\rm Pia}} = p_{\rm CSF} \}
\end{align}
To discretize~\eqref{eq:1d_flow} with the bifurcation
conditions~\eqref{eq:1d_flow:bifurcations} and boundary
conditions~\eqref{eq:1d_flow:bcs}, we use the space of continuous
functions $\hat{Z}_h(\mathcal{I})$ to enforce the continuity of the pressure at
bifurcation points, while the conservation of flux is enforced
(weakly) through the variational formulation. The discrete variational form of the equations then reads: Find $(\hat{q}, \hat{p}) \in \hat{Z}_h(\mathcal{I}) \times \hat{V}_{h, p_0}(\mathcal{I})$ such that
\begin{subequations}
\begin{align}
  \langle \hat{q}, z \rangle +  \langle z, \frac{A \kappa}{\mu} \partial_s \hat{p}  \rangle &= 0 \quad \forall z \in \hat{Z}_h(\mathcal{I}) \\
  \langle \hat{q}, \partial_s w \rangle &= 0 \quad  \forall w \in \hat{V}_{h, 0}(\mathcal{I}).
\end{align}
\end{subequations}%
where $\langle \cdot, \cdot \rangle$ denotes the $L^2(\mathcal{I})$-inner product and is defined segment-wise. For a stability and convergence analysis of this discrete model, we refer to the reference\cite{gjerde2024directional}.

@@

\subsection{Extended model validation}
\label{sec:app:model_validation}

>>>>

\renewcommand\thefigure{S\arabic{figure}}    
\renewcommand\thesection{S\arabic{section}}    
\renewcommand\thetable{S\arabic{table}}

@@

These sections provide more detailed descriptions of the mathematical
models and numerical approximations considered.

@@

In addition, we consider two sets of perivascular networks: a
periarterial network $\Lambda_a$ represented by the (connected)
centerlines $\Lambda^i_{a}$ of the arterial tree, and a perivenous
network $\Lambda_v$ associated with the centerlines $\Lambda^i_{v}$ of
the veins. Recall that we represent the vascular domains as the union
of cylindrical vessels of radius $R_1^i$ surrounding the centerlines
$\Lambda^i$. Moreover, we consider the surrounding perivascular spaces
as the union of annular cylinders $\Omega^i$ of inner radius $R_1^i$
and outer radius $R_2^i > R_1^i$ and thus of width $R_2^i - R_1^i$,
and remark that we interpret $R^i_2$ and $R^i_1$ as temporal averages
(fixed in time), as the pulsations considered in
\Cref{sec:app:peristalsis} are beyond the temporal resolution of our
model. We denote the outer lateral surface of the periarterial and
perivenous spaces by $\Gamma_a$ and $\Gamma_v$, respectively. We omit
the subscript $a$ or $v$ when referring to any such network, vessel
segment or perivascular outer surface. We assume that $\Lambda$ is
parametrized by the coordinate $s$, and with a minor abuse of
notation, simply write $s \in \Lambda$ to represent the point ${\bm
  \lambda}(s)$ on $\Lambda$ corresponding to $s$. Finally, we view
each perivascular network both as a geometric domain and as a directed
graph with the centerlines $\{ \Lambda^i \}$ as oriented edges and the
connections as nodes $\mathcal{V}$. The notation used for the
geometry, model variables and parameters is summarized in
Table~\ref{table:geometry_notation} and \ref{table:variable}.
\begin{table}[h!] 
\centering
\begin{tabular}{ll}
\toprule
\textbf{Symbol} & \textbf{Description} \\
\midrule
$\Omega$ & Full intracranial domain: $\Omega = \Omega_{\mathrm{PAR}} \cup \Omega_{\mathrm{CSF}}$ \\
$\Omega_{\rm PAR}$ & Brain parenchyma domain \\
$\Omega_{\mathrm{CSF}}$ & Cerebrospinal fluid (CSF) domain \\
$\partial\Omega_{\mathrm{PAR}}\cap\partial\Omega_{\mathrm{CSF}}$ & Parenchyma–CSF interface: $\partial \Omega_{\rm CSF} \bigcap \partial
\Omega_{\rm PAR} = \Gamma_{\rm LV} \cup \Gamma_{\rm pia}$ \\
$\Gamma_{\mathrm{LV}}$ & Lateral ventricle surface (part of parenchyma–CSF interface) \\
$\Gamma_{\mathrm{pia}}$ & Pial surface (remainder of parenchyma–CSF interface) \\
$\Gamma_{\mathrm{SSAS}}$ & Lower CSF boundary toward spinal subarachnoid space \\
$\Gamma_{\mathrm{AM}}$ & Outer CSF boundary toward arachnoid/dura \\
$\Gamma_{\mathrm{AM\text{-}L}}$, $\Gamma_{\mathrm{AM\text{-}U}}$ & Lower/upper parts of $\Gamma_{\mathrm{AM}}$ \\
$\Gamma_{\mathrm{SC}}$ & Parenchyma–spinal‐cord boundary: $\partial\Omega_{\mathrm{PAR}}\setminus\partial\Omega_{\mathrm{CSF}}$ \\
$\Lambda_{a},\Lambda_{v}$ & Arterial/venous PVS networks \\
$\Lambda^{i}$ & Individual PVS centerline segment \\
$L_i$ & Length of a PVS segment $\Lambda^i$ \\
$R_{1}^{i},\,R_{2}^{i}$ & Inner/outer radii of the annular PVS \\
$\Gamma_{a},\,\Gamma_{v}$ & Outer lateral surfaces of periarterial/perivenous PVSs \\
$A^i$ & Area of the cross‐section for PVS segment $\Lambda^i$ \\
$P^i$ & Perimeter of the cross‐section for PVS segment $\Lambda^i$ \\
$\mathcal V$ & Graph nodes of the PVS network \\
$\mathcal B\subset  \mathcal V$ & Internal (bifurcation) nodes \\
$\mathcal N\subset  \mathcal V$ & End (leaf) nodes \\
\bottomrule
\end{tabular}
\caption{Geometrical notation, refer to (Fig.~1A, Fig.~2A--C)}
\label{table:geometry_notation}
\end{table}
\begin{table}[htb!]
\centering 
\begin{tabular}{ll}
\toprule
\textbf{Symbol} & \textbf{Description} \\
\midrule
$\phi$ & Fluid volume fraction (porosity) \\
$D$ & Effective diffusion coefficient ($D_{\mathrm{PAR}} = D^*$ or $D_{\mathrm{CSF}} = D(1+R)$ \\
$\bm u$ & Convective velocity (${\bm u}_{\mathrm{PAR}} = 0$ or ${\bm u}_{\mathrm{CSF}}$ computed from steady state flow) \\
$c$ & Solute concentration in 3D domains \\
$\overline{c}$ & Lateral average of $c$ on the PVS boundary \\
$\xi$ & Permeability coefficient between the periarterial and perivenous networks
and their surroundings\\
$\delta_{\Gamma}$ & Dirac term on the PVS outer surfaces \\
$\beta_{\mathrm{pia}}$ & Permeability of pial/ventricle interface \\
$\beta_{\mathrm{exit}}$ & Efflux rate on $\Gamma_{\mathrm{AM\text{-}U}}$ \\
$g_{\mathrm{influx}}$ & Influx on $\Gamma_{\mathrm{SSAS}}$ \\
$\hat D$ & Effective diffusion coefficient in the PVS networks \\
$\hat u$ & Convective velocity in PVS networks \\
$p$, $\hat{p}$ & Pressures in the CSF spaces and PVS networks \\
$\hat{q}$ & Volumetric flux in PVS network \\
$\kappa$ & Hydraulic conductivity in annular PVS \\
$\mu$ & Dynamic viscosity of CSF \\
$R_{\rm CSF, 0}$ & Resistance parameter for CSF efflux \\
$u_{\mathrm{in}}$ & Imposed fluid influx on $\Gamma_{\mathrm{LV}}$ \\
$\omega,\;\rho$ & Angular frequency, CSF density (for dispersion estimates) \\
$\alpha_c,\, \alpha_r$ & Womersley number (cardiac, respiratory) \\
$h_{\rm SAS}$ & Mean width of the subarachnoid space (for dispersion estimates) \\
$R_c,\,R_r$ & Spatially-varying dispersion enhancement factors (cardiac, respiratory) \\
\bottomrule
\end{tabular}
\caption{Variable and parameter notation}
\label{table:variable}
\end{table}

%% \begin{table}[h!]
%% \centering
%% \caption{Common mesh and finite element notation}
%% \label{table:mesh}
%% \begin{tabular}{ll}
%% \hline
%% \textbf{Symbol} & \textbf{Description} \\
%% \hline
%% $\mathcal{T}$ & Tetrahedral mesh of $\Omega$ \\
%% $\mathcal{T}_{\mathrm{PAR}},\,\mathcal{T}_{\mathrm{CSF}}$ & Restrictions of $\mathcal T$ to $\Omega_{\mathrm{PAR}},\Omega_{\mathrm{CSF}}$ \\
%% $E$ & Tetrahedral cell in $\mathcal T$ \\
%% $\mathcal F_{i}$ & Interior facets of $\mathcal T$ \\
%% $\mathcal F_{i,\mathrm{PAR}},\, \mathcal F_{i,\mathrm{CSF}}$ & Interior facets in $\Omega_{\mathrm{PAR}}$, $\Omega_{\mathrm{CSF}}$ \\
%% $\mathcal F_{e}$ & Exterior facets of CSF mesh ($\mathcal F_{\mathrm{pia}}\cup \mathcal F_{\mathrm{LV}}\cup \mathcal F_{\mathrm{AM\text{-}L}}\cup \mathcal \mathcal F_{\mathrm{AM\text{-}U}}\cup \mathcal F_{\mathrm{SSAS}}$) \\
%% $\mathcal F _{\mathrm{pia}},\,\mathcal F_{\mathrm{LV}},\, \mathcal F_{\mathrm{AM\text{-}L}},\,\mathcal F_{\mathrm{AM\text{-}U}},\, \mathcal F_{\mathrm{SSAS}}$ & Facets on corresponding boundaries \\
%% $V_{h}(\mathcal{T})$ & Space of piecewise linear polynomials on $\mathcal{T}$\\
%% $Q_{h}(\mathcal{T})$ & Pressure space on $\mathcal T_{\mathrm{CSF}}$  \\
%% $\mathcal{I}_a, \mathcal{I}_v$ & 1D meshes of arterial/venous networks $\Lambda_{a},\Lambda_{v}$ \\
%% $\widehat V_{h}(\mathcal{I})$ & Space of continuous piecewise linear polynomials on $\mathcal{I}$, $\mathcal{I}$ omitted when superfluous \\
%% $\widehat{Z}_{h}(\mathcal{I})$ & Space of piecewise constants ($P^0$) on $\mathcal{I}$, , $\mathcal{I}$ omitted when superfluous \\% $\widehat V_{h,0}$, $\widehat V_{h,g}$ & Continuous Galerkin space on network with homogeneous or prescribed BCs
%% $\mathrm{BDM}_{2}$ & Brezzi–Douglas–Marini space of degree 2 \\
%% $\tau$ & Time step size \\
%% \hline
%% \end{tabular}
%% \end{table}

@@

\subsubsection{Finite element solution of the incompressible Stokes equations}
We consider an H(div)-based finite element approximation of the incompressible Stokes equations~\eqref{eq:stokes} defined over $\Omega_{\rm CSF}$ with the boundary conditions~\eqref{eq:stokes_bcs}. Following\cite{hong2016robust}, we approximate the velocity field $\bm u$ and the pressure field $p$ with the following finite element spaces:
\begin{align*}
  \bm V_{h,g} &= \{ \bm v  \in H(\mathrm{div}, \Omega_{\mathrm{CSF}}): \;
  \bm v \vert_{E} \in \mathrm{BDM}^2(E), \; E \in \mathcal{T}_{\rm CSF}; \;
  \bm v \cdot \bm n = 0
  \text{ on } \Gamma_{\rm{pia}} \cup \Gamma_{\rm AM-L} \cup \Gamma_{\rm{SSAS}}, \;\; \bm v \cdot \bm n =  g %\frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}} 
  \text{ on } \Gamma_{\mathrm{LV}} \}  \\ 
  Q_h  &= \{q \in L^2(\Omega_{\rm CSF}): \; q \vert_E \in \mathrm{P}^1(E),
  \; E \in \mathcal{T}_{\rm CSF} \}. 
\end{align*}
Here, $H(\mathrm{div}, \Omega_{\rm{CSF}})$ is the space of $L^2(\Omega_{\rm{CSF}})$ vector fields with $L^2(\Omega_{\rm{CSF}})$ divergence, $\mathrm{BDM}^2$ is the Brezzi-Douglas--Marini element~\cite{brezzi1987mixed} of degree 2, $g$ is a given constant, and $\bm n$ is the unit outward normal vector to each facet. Given any vector $\bm v$, the normal and tangential components on each facet are denoted and given by 
$$ \bm v_n = (\bm v \cdot \bm n) \bm n, \quad \bm v_t = \bm v - \bm v_n. 
$$ 
Since $\bm V_{h,g} \subset H(\mathrm{div}, \Omega_{\rm CSF})$, then $[\bm
  v_n] = 0 $ on $\mathcal{F}_{i, \rm CSF}$, the interior facets to $\Omega_{\mathrm{CSF}}$. Continuity in the
tangential component is enforced weakly via interior penalization. For
convenience, we collect all facets exterior to the CSF space  $\Gamma_{\rm AM-U}$ in the set
$$\mathcal{F}_{e} = \mathcal{F}_{\rm pia} \cup \mathcal{F}_{\rm LV}
\cup \mathcal{F}_{\rm AM-L}\cup \mathcal{F}_{\rm AM-U}  \cup \mathcal{F}_{\rm SSAS}, $$ 
where facets
lying on the pial interface $\Gamma_{\rm pia}$ are denoted by
$\mathcal{F}_{\rm pia}$, on the lower and upper outer (arachnoid) boundary
$\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$ by $\mathcal{F}_{\rm AM-L}$ and $\mathcal{F}_{\rm AM-U}$, respectively, 
%on the boundary toward the spinal cord $\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$
 on the boundary towards the spinal SAS $\Gamma_{\rm SSAS}$ by
$\mathcal{F}_{\rm SSAS}$, and on the surface of the lateral ventricles by $\mathcal{F}_{\rm LV}$. Now, define the form 
\begin{multline}
  \mathcal{A}_h(\bm u, \bm v)
  = \sum_{E \in \mathcal{T}_{\rm CSF}} \int_{E} \mu \nabla \bm u : \nabla \bm v
 + \sum_{F \in \mathcal{F}_{i, \rm CSF} \cup \mathcal{F}_{e}} \left (
  -\int_{F} \mu \{\nabla \bm u\}  \bm n_F \cdot [\bm v_t] 
  - %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
  \int_{F} \mu \{\nabla \bm v\}  \bm n_F \cdot [\bm u_t]
  + %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
\int_{F}   \frac{\sigma \mu}{h_F} [\bm u_t ] \cdot [\bm v_t] \right ),   
\end{multline}
where on exterior facets the average and jump operators take the one-sided values. We set the penalty parameter for the tangential continuity to be $\sigma = 20$. The finite element discretization of the incompressible Stokes equations is then to find $(\bm u_h, p_h) \in \bm V_{h,g} \times Q_h$ with $g  = \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}}$  such that 
\begin{subequations}
\begin{align}
  \mathcal{A}_h(\bm u_h, \bm v_h)
  + \sum_{F \in \mathcal{F}_{\rm AM-U}} \int_{F} R_{\rm CSF, 0} (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n)
  - \int_{\Omega_{\rm CSF}} \nabla \cdot \bm v_h \, p_h  &= 0 \qquad \foralls \bm v_h \in \bm V_{h,0} \\ 
\int_{\Omega_{\rm CSF}} \nabla \cdot \bm u_h \, q_h  &= 0 \qquad \foralls q_h \in Q_h.
\end{align}
\end{subequations}

@@

\subsubsection{Finite element solution of the perivascular network equations} 
We consider meshes $\mathcal{I}_a, \mathcal{I}_v$ representing a
conforming subdivision of each of the perivascular networks
$\Lambda_a$, $\Lambda_v$. Relative to each mesh $\mathcal{I}$, we define the
space of (discontinuous) piecewise constants $\widehat{Z}_h(\mathcal{I})$ and define
spaces of continuous piecewise linears $\hat{V}_{h, g}(\mathcal{I})$ with
prescribed boundary node values (on $\mathcal{N}$) given by $g$:
\begin{align}
  \hat{Z}_h(\mathcal{I}) &= \{z \in L^2(\Lambda) ; \, z \vert_{\Lambda^i} \in \mathrm{P}^{0}(\Lambda^i) \, \foralls \Lambda^i \in \mathcal{I}\}, \\
  \hat{V}_{h, g}(\mathcal{I}) &= \{ v \in C^0(\Lambda) ; \, v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda^i), \, v(x) = g(x) \, \foralls x \in \mathcal{N} \} .
  %Q_h^{\rm PVS} &= \{ q_h \in C^0(\Lambda),  q_h \vert_{\Lambda^i} \in \mathrm{P}^k(\Lambda^i), q_h \vert_{\Gamma_{\rm Pia}} = p_{\rm CSF} \}
\end{align}
To discretize~\eqref{eq:1d_flow} with the bifurcation
conditions~\eqref{eq:1d_flow:bifurcations} and boundary
conditions~\eqref{eq:1d_flow:bcs}, we use the space of continuous
functions $\hat{Z}_h(\mathcal{I})$ to enforce the continuity of the pressure at
bifurcation points, while the conservation of flux is enforced
(weakly) through the variational formulation. The discrete variational form of the equations then reads: Find $(\hat{q}, \hat{p}) \in \hat{Z}_h(\mathcal{I}) \times \hat{V}_{h, p_0}(\mathcal{I})$ such that
\begin{subequations}
\begin{align}
  \langle \hat{q}, z \rangle +  \langle z, \frac{A \kappa}{\mu} \partial_s \hat{p}  \rangle &= 0 \quad \forall z \in \hat{Z}_h(\mathcal{I}) \\
  \langle \hat{q}, \partial_s w \rangle &= 0 \quad  \forall w \in \hat{V}_{h, 0}(\mathcal{I}).
\end{align}
\end{subequations}%
where $\langle \cdot, \cdot \rangle$ denotes the $L^2(\mathcal{I})$-inner product and is defined segment-wise. For a stability and convergence analysis of this discrete model, we refer to previously presented results\cite{gjerde2024directional}.

@@

\subsection{Extended comparison with literature}
\label{sec:app:model_validation}
